---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(tidyverse)
Valores_DistCosta <- readRDS("/home/giorgia/Documents/Doctorado tesis/Mod.distrib/ModelacionAsentamientosGit/DistCosta_Chile.rds")
Indice <- readRDS("/home/giorgia/Documents/Doctorado tesis/Mod.distrib/ModelacionAsentamientosGit/Indice.rds") %>% rename(Ciudad= City)
Validacion_Indice <- left_join(Indice, Valores_DistCosta)

Validacion_Indice <- Validacion_Indice %>% dplyr::filter(!is.na(Dist_Costa)) %>% rename(City=Ciudad)

#Validacion_IndiceSur <- Validacion_Indice %>% dplyr::filter((Lat  <= -40 ))
#Validacion_IndiceNorte <- Validacion_Indice %>% dplyr::filter((Lat  >= -40 ))
#Validacion_IndiceValpo <- Validacion_Indice %>% dplyr::filter(Ciudad %in% c("Valparaíso", "Viña del Mar", "Concón", "Gran Valparaíso"))

```

# MarineIndex


The goal of MarineIndex is to develop a index of people connections with the sea.
In **Idea.R** we extract information of Wikipedia as peoples's perceptions. With the code we detected all cities from Chile and we made an scraping to extract all the words of each city's text. 

An app was created to input values to all words detected: marine +1, terrestrial -1, and neutral words 0. See **TestApp**.

In **Relaciones indice marino.R** we related marine index with possible explanatory variables. We used *DistCosta_Chile.rds* and *Indice.rds*.



```{r}



ggplot(Validacion_Indice) + geom_point(aes(x=Dist_Costa, y=Index)) +  scale_colour_continuous()+
  ylab("Indice conexión marina")+ xlab("Distancia a la costa (metros)")+ 
  scale_x_continuous(label = scales::comma)+ theme_classic()

```

This code is feed by variables created in other scrips:

- **DiversidadAves.R**: bird diversity for each Chilean city with data obtain from iNaturalist. We created *Bird_type_prop.rds*.

- **Viewshed_altitud.R** created a value of coastal visibility from 20% of the cells of each city.


You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this. You could also use GitHub Actions to re-render `README.Rmd` every time you push. An example workflow can be found here: <https://github.com/r-lib/actions/tree/v1/examples>.






In that case, don't forget to commit and push the resulting figure files, so they display on GitHub.
